### 2018-01-10 ident/docs/ident_manuscript.tex
updates to intro

If all intracellular metabolite concentrations can be measured over a time course, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}), based on the measured data. The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations can be used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}), with fixed upper ($\theta_u$) and lower ($\theta_l$) bounds for the parameter values (Equation \ref{eq:lsqopt}b). The difference between the data and the model estimate for each output and at each time point in the objective function is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
\begin{subequations}\label{eq:lsqopt}
	\begin{align}
	\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{x_{kl}^*-x_{kl}}{\sigma_{kl}^*}\right)^2\\
	&\theta_l \le \theta \le \theta_u
	\end{align}
\end{subequations}

However, not all metabolite concentrations used in the model (Equation \ref{eq:kinstoich}) can be measured. Additionally, measurable fluxes in the metabolic network also need to be included as part of the parameter estimation problem. In such scenarios, the parameter estimation problem is modified to suit a new system of equations shown below (Equation \ref{eq:output}). The new system of equations is obtained by augmenting the original system (Equation \ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations and fluxes ($y$) and the unmeasured concentrations ($x$) that are used in the original model (Equation \ref{eq:kinstoich}) above. Equation (\ref{eq:output}) now has additional parameters $\mu$ that define this relationship and also need to be estimated.
\begin{subequations}\label{eq:output}
	\begin{align}
	\dot{x} = \mathbf{S}v\\
	v = f(x, y, \theta, u)\\
	\dot{y} = h(x, y, \theta, u)
	\end{align}
\end{subequations}
In systems identification, the measured concentrations and fluxes ($y$) are called output or observed variables, and the unmeasured concentrations ($x$) are called the state variables. For estimating $\theta$, the metabolite concentrations $x$ in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.

@@

Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems \parencite{Audoly2001}(IEEE Trans paper from 2007). Most of these algorithms use methods based on differential algebra \parencite{Ljung1994}. More recent methods take a profile-likelihood-based approach to establish both structural and practical identifiability \parencite{Raue2009a}. All of these methods are heavily dependent on the availability of dynamic time course data for all the output variables of the system and scale poorly with increase in size of the modeled system \parencite{Bellu2007}.

Computational burden due to poor scalability can be partly addressed with the current increases in computational power. Although, improvements in measurement technologies can enable one to obtain high frequency dynamic data, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.

>>>>

If all intracellular metabolite concentrations can be measured over a time course, a nonlinear programming problem can be formulated to estimate the enzyme kinetic parameters ($\theta$) in Equation (\ref{eq:kinstoich}), based on the measured data. The minimization of least square error between the measured ($x^*$) and modeled ($x$) concentrations is used as an objective function (Equation \ref{eq:lsqopt}a) for the optimization problem (Equation \ref{eq:lsqopt}), with fixed upper ($\theta_u$) and lower ($\theta_l$) bounds for the parameter values (Equation \ref{eq:lsqopt}b). The difference between the data and the model estimate for each output and at each time point in the objective function is weighted by the variance in the experimental data $\sigma_{kl}^*$ for that corresponding variable and time point.
\begin{subequations}\label{eq:lsqopt}
	\begin{align}
	\underset{\theta}{\mathrm{min}} &\sum_{k=1}^{m}\sum_{l=1}^{d}\left(\frac{x_{kl}^*-x_{kl}}{\sigma_{kl}^*}\right)^2\\
	&\theta_l \le \theta \le \theta_u
	\end{align}
\end{subequations}

However, not all metabolite concentrations used in the model (Equation \ref{eq:kinstoich}) can be measured. Additionally, measurable fluxes in the metabolic network also need to be included as part of the parameter estimation problem. In such scenarios, the parameter estimation problem is modified to suit a new system of equations shown below (Equation \ref{eq:output}). The new system of equations is obtained by augmenting the original system (Equation \ref{eq:kinstoich}) with Equation (\ref{eq:output}c) that models the relationship between the measurable metabolite concentrations and fluxes ($y$) and the unmeasured concentrations ($x$) that are used in the original model (Equation \ref{eq:kinstoich}) above. The parameter vector ($\theta$) is augmented with additional parameters that define this relationship. These additional parameters also need to be estimated.
\begin{subequations}\label{eq:output}
	\begin{align}
	\dot{x} = \mathbf{S}v\\
	v = f(x, y, \theta, u)\\
	\dot{y} = h(x, y, \theta, u)
	\end{align}
\end{subequations}
In systems identification, the measured concentrations and fluxes ($y$) are called output or observed variables, and the unmeasured concentrations ($x$) are called the state variables. For estimating $\theta$, the metabolite concentrations $x$ in the optimization problem (Equation \ref{eq:lsqopt}) are substituted with the output variables $y$.

@@

Algorithms for structural identification of parameters have 
Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems \parencite{Audoly2001}(IEEE Trans paper from 2007). Most of these algorithms use methods based on differential algebra \parencite{Ljung1994}. The underlying motivation for these methods is the need to change the non-convex parameter estimation problem (Equation \ref{eq:lsqopt}) to a convex problem due to inherent difficulties in obtaining global optimal solutions for non-convex problems \parencite{Bellu2007}. However, due to the use of computer algebra systems to solve the identifiability problem, these methods are not scalable to large systems like a kinetic model of a large metabolic network .

Recently proposed methods address the issue of scalability by using a profile-likelihood-based approach to establish structural identifiability \parencite{Raue2009a}. Profile likelihood-based methods offer an advantage over the differential algebra methods due to their ability to not only structurally identify parameters, but also help with experimental design through practical identifiability of parameters.

In contrast to differential algebra-based methods, these methods still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters. Hence, these methods can still be considered to be un-scalable for structurally identifying parameters in large metabolic networks. Nonetheless,

Apart from issues with their scalability, profile likelihood-based methods also depend on the availability of actual measured or simulated dynamic data to establish structural identifiability.

Although, improvements in measurement technologies can enable one to obtain high frequency dynamic data to address the need for time course data required by these methods,

The profile likelihood-based methods and the differential algebra-based methods are heavily dependent on the availability of

 for all the output variables of the system and scale poorly with increase in size of the modeled system.

However, computational burden due to poor scalability can be partly addressed with the current increases in computational power.

, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.
